### 2023-10-15 Paper/Intro.Rmd
Separate file with only intro

---
title: "Learning with Misspecified Models: the case of overestimation"  
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Jimena Galindo"
abstract: "I design a framework and a laboratory experiment that allows for the comparison of multiple theories of misspecified learning. I focus on a framework with endogenous information and a data-generating process ruled by two fundamentals: an ego-relevant parameter and a state. Within this framework I study three forces that can lead to misspecified beliefs: initial misspecifications, learning traps and biased updating. I find that biased updating is the main driver of misspecified beliefs in the lab. In addition, I vary the degree of ego-relevance of the parameter by introducing a stereotype treatment. The data is consistent with biased updating in both cases but for potentially different reasons: when learning about themselves, subjects attribute successes to their own ability and failures to luck. Instead, in the stereotype treatment, they compensate for initial negative biases by over-attributing positive signals to the ability of others. This translates into similar observed choices but different dynamics in beliefs."

#classoption: pagebackref # passes pagebackref=true to hyperref before it is loaded, allowing backreferencing
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7  
        fig_height: 6
        fig_caption: true
        number_sections: true

        template: NULL
        keep_tex: true
        extra_dependencies:
            footmisc: ["bottom"] # footnote management 
            setspace: ["doublespacing"] # spacing of the paper 
            caption: ["normal"] 
            dsfont: null # indicator function 1
            booktabs: null 
            makecell: null 
            hyperref: null

        includes:
          in_header: extra_header.tex

bibliography: references.bib
fontsize: 12pt 
geometry: margin=2.5cm

---

```{r dependencies, include=FALSE}
source("../Paper/parameters_and_packages.R")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 7, fig.height = 6, fig.align = "center")
```

```{r data, include=FALSE}
updates <- fread("../Clean/updates_post.csv")
```

\newpage
# Introduction

A growing body of literature in economics explores how people develop incorrect beliefs 
about fundamentals. Most of this research centers on scenarios where agents 
passively observe the world, and incorrectly integrate the received information into their 
beliefs.\footnote{See \citet{benjamin2019} for a review of the literature on errors of 
probabilistic thinking.}
However, many real-world situations cast agents as active participants in the generation of 
information. In these cases, the information they observe is influenced by their actions and 
subsequent behavior is in turn determined by how they incorporate the information 
into their beliefs.

As an example consider a student who needs to decide how much effort to put into 
studying for an exam. Their decision will depend on two factors: their belief about their intrinsic 
ability, and their belief about how difficult the exam will be. The outcome 
they observe will be affected by how much they decide to study. Imagine that the student 
puts in a moderate amount of effort and gets a surprisingly good grade. Did they get a good grade because 
they are smarter than they thought? Or because the exam was easier than they had anticipated? 
Their future exam preparation strategy will depend fundamentally on which line of reasoning they 
take. This feedback loop is referred to as an \emph{endogenous information 
process} and is at the center of the forces I study.

To understand what the main forces at play are, I compare a set of theories that model learning in settings with 
endogenous information, and which can rationalize the persistence of misspecified beliefs. I develop a unifying framework that nests multiple 
theories of learning and generates testable predictions for each of them. 
Then I cast this framework in a laboratory experiment and test the predictions to identify which 
of the theories are consistent with the behavior I observe in the lab.
The experiment features an agent who needs to 
learn two parameters: one that pertains their own characteristics 
(an \emph{ego-relevant} parameter), 
and an exogenously determined state---in the context of the student preparing for exams, 
the two parameters  would be their intrinsic ability and the difficulty of the exam.

Misspecified beliefs about an ego-relevant parameter are often referred to as overconfidence 
(or underconfidence) and 
have been documented by behavioral scientists\footnote{\citet{kelley1980} 
provides a review of the psychology literature.} and economists\footnote{See \citet{benjamin2019} 
for a review of the literature in economics.} in a variety of settings.  @Oster2013 find that 
subjects who are at high risk of having Huntington's disease overestimate their probability of 
being healthy and make retirement decisions as if they were healthy. @Hoffman2020 show that workers 
overestimate the quality of their match to their 
current employment and are unlikely to look for other opportunities. @Camerer1999 find that 
entrepreneurs are overconfident about the quality of their enterprise, 
which leads to excessive entry and early exit from markets.

In all of these examples,
holding an incorrect belief about a fundamental leads to suboptimal choices with 
potentially high costs. In spite of the abundance of evidence, the scope of the existing research 
is limited in terms of the frameworks it considers. Most of the experimental evidence 
documenting the bias is collected in settings where subject are passive learners.\footnote{
\citet{Gotte2022} and \citet{Ozyilmaz2022} are exceptions that study settings with endogenous information processes.} They 
observe a noisy signal and report their beliefs over subsequent rounds.\footnote{\citet{Bracha2012} and 
\citet{Mobius2022} are some examples.} Although these studies provide important insights, they are 
not flexible enough 
to incorporate the richer theories that have been proposed more recently. In particular, they 
do not allow the study of endogenous learning or for learning about multiple parameters at once.\footnote{\citet{Coutts2020} 
studies an environment with an ego-relevant parameter and an exogenous state but does not 
incorporate the endogenous information process.}

In my experiment, I move away from the standard framework of passive learning to analyze a richer set of learning mechanisms. In particular, 
the interaction between the two parameters together with an endogenous information process 
gives rise to three possible mechanisms that allow
for the persistence of incorrect beliefs: the presence of learning traps, incorrect initial beliefs, and misattribution bias. The theories that I consider incorporate different combinations of these mechanisms.

When the setting features learning traps, even an agent who incorporates all information correctly, 
may fall into learning traps as outlined by @Hestermann2021. These traps are characterized by a 
combination of an incorrect belief and an
optimal action which produce information that confirms the incorrect belief. 
Once an agent falls into a trap, the belief will be stable and even with 
a correctly specified model of learning, they will not be able to abandon their misspecified beliefs. 
If, the agent is 
dogmatic about their initial belief, @Heidhues2018 show that they will inevitably fall into a trap
and thus will be able to rationalize and sustain their initial misspecification belief.\footnote{\citet{Gotte2022} study the case of agents with dogmatic initial beliefs in a 
laboratory experiment.}

@Ba2023 moves away from dogmatism and endows the agents with a mechanism through which 
they can abandon incorrect beliefs; this allows them to avoid falling into learning traps. 
To do so agents perform Bayesian hypothesis tests that evaluate which is the more 
likely parameter out of two possibilities. By doing this, she characterizes the set of situations 
in which, even agents who consider alternative paradigms, may become trapped.\footnote{A 
similar mechanism is proposed by \citet{Schwarstein2021} in a setting with persuasion.}

Lastly, misattribution bias is the more classical explanation and has been widely studied in 
behavioral science.\footnote{See \citet{kelley1980} for a review.}
Agents who suffer from misattribution bias will attribute successes to their own ability---the 
ego-relevant parameter---and failures to bad luck---the state. 
Under this model of learning, even an agent who initially has a correct initial 
belief may become overconfident if they observe a sequence of successes. In this case, 
the main driver of the bias is not an initial misspecification or the presence of learning traps, 
it is the updating procedure itself.\footnote{A more general framework that can be used to model 
this bias has also been proposed by \citet{Brunnermeier2005} and empirically studied by \citet{Bracha2012}.}

These theories provide the main building blocks for a simplified framework that can be directly 
implemented in a laboratory experiment. In the experiment subjects make choices and receive feedback that depends on their own ability, an exogenous parameter and the choice they made. I track their choices as well as their beliefs about their own ability. The goal is to 
identify which of the 3 forces---the presence of learning traps, misspecified initial beliefs, or misatribution 
bias---better explains the observed behavior. To determine the fit of the models, I compare the behavior
predicted by each theory to the benchmark given by the fully Bayesian updating procedure.

I also study whether the learning mechanism is inherently linked to the ego-relevance of the parameters 
or if it is a more general phenomenon. I vary the degree of ego-relevance by introducing a 
treatment in which subjects learn about the ability of another participant.
In this treatment, the participants know only the gender and nationality of the other, and thus can 
induce stereotypes---a different type of misspecification.  
If correct learning about the parameters happens at higher rates in the stereotype treatment, 
it would suggest that the 
bias is intrinsically linked to the ego-relevance of the parameter. In contrast, if similar 
biased behavior arises in both treatments, it is more likely that the main driver of these types 
of misspecified beliefs is the updating procedure itself or the endogenous information process.

Although some agents do fall into learning traps, I find that the behavior of most 
subjects is better explained by misattribution bias: good news are treated as signaling high ability, while bad news are attributed to a low state. I also find that misattribution is 
no more prevalent in the ego-relevant than in the stereotype condition. This suggests 
that the main driver of the misspecification is the updating procedure; however, the underlying 
mechanism by which the bias is generated may be 
different in both treatments: While in the ego-relevant condition subjects prefer to hold themselves in high esteem, in the stereotype condition
updating seems to be driven by some sort of bias overcorrection---when subjects realize that they underestimated 
the ability of another participant based on their gender and nationality, they compensate by
overestimating their ability.

Finally, I estimate the structural parameters of the models to study model-heterogeneity in the sample.
I find that even at an individual level the behavior is better explained by a general model of 
misattribution bias for most
subjects. There is a smaller group of subjects that can be better explained through dogmatic beliefs
and hypothesis testing and none of them behave in line with the fully Bayesian benchmark.

In what follows I first discuss the theoretical framework and the predictions of each of the theories.
Then I introduce a unifying example and my hypotheses. In section 4 I describe the experimental
design and in section 5 I present the data and the results. Section 6 outlines the estimation of the parameters
and the model fit analysis.